# Nodal Density in Penile Cancer

__Title:__ High Nodal Density Is Associated with Aggressive Pathologic Features and Predicts Outcome in Patients with Penile Squamous Cell Carcinoma Treated by Penectomy and Inguinal Lymphadenectomy

__List of Authors:__ Isabel Alvarado,<sup>1</sup>  César Aponte,<sup>2</sup>, Diana Piedras,<sup>1</sup> Adriana Rodriguez-Gómez,<sup>1</sup> Diego F Sanchez,<sup>3</sup> Narciso Hernández-Toriz,<sup>1</sup> Antonio L. Cubilla,<sup>3</sup> and  Alcides Chaux<sup>2</sup>

<sup>1</sup> Servicio de Anatomía Patológica, Hospital de Oncología, Centro Médico Nacional Siglo XXI, Instituto Mexicano del Seguro Social, México D.F.  
<sup>2</sup> Office of Scientific Research, Norte University, Asunción, Paraguay  
<sup>3</sup> Instituto de Patología e Investigación, Facultad de Ciencias Medicas,  Asunción, Paraguay

***

Last update: ```r {date()}```

```{r GlobalOptions}
# Setting the global options for the analysis
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r DataSet}
source("StandardAnalysis3.R")
library(foreign)
library(car) # For scatterplot
library(effsize) # For cohen.d
# Reading the full dataset
DataFull <- read.dta("mx_penile_dataset_february_2014.dta")
# Creating a list of selected variables
SelectedVariables <- c("age", "penectomy", "site", "size", "subtype", "grade", "thick", "lvi", "pni", "level", "ln_dx", "ln_r", "r_met", "ln_l", "l_met", "ln_t", "met_dens", "ln_met", "met_bi", "rt_pos", "outcome", "cancer_dis", "cancer_death", "hpvrelated", "pi", "pt_stage", "pn_stage", "m_stage", "tnm_stage", "urethra", "fu_month")
# Creating the working dataset
Data <- DataFull[, SelectedVariables]
Data <- droplevels(Data)
# Renaming levels of factors
library(plyr)
Data$site <- revalue(Data$site, c(
        "C" = "Coronal Sulcus",
        "F+C" = "Foreskin + Coronal Sulcus",
        "G" = "Glans",
        "G+C" = "Glans + Coronal Sulcus",
        "G+F" = "Glans + Foreskin",
        "G+S" = "Glans + Shaft"
        ))
Data$level <- revalue(Data$level, c(
        "LP" = "Lamina Propria",
        "CS" = "Corpus Spongiosum",
        "CC" = "Corpus Cavernosum"
        ))
Data$grade <- revalue(as.factor(Data$grade), c(
        "1" = "Grade 1",
        "2" = "Grade 2",
        "3" = "Grade 3"
        ))
Data$subtype2 <- revalue(Data$subtype, c(
        "Verrucous-Papillary-Warty" = "Mixed W/B",
        "Verrucous-Papillary" = "Mixed non-W/B",
        "Usual-Warty-Basaloid" = "Mixed W/B",
        "Usual-Warty" = "Mixed W/B",
        "Usual-Verrucous-Warty" = "Mixed W/B",
        "Usual-Verrucous-Papillary" = "Mixed non-W/B",
        "Usual-Verrucous-Cuniculatum" = "Mixed non-W/B",
        "Usual-Verrucous" = "Mixed non-W/B",
        "Usual-Papillary-Warty" = "Mixed W/B",
        "Usual-Papillary" = "Mixed non-W/B",
        "Usual-Basaloid" = "Mixed W/B"
        ))
Data$site2 <- revalue(Data$site, c(
        "Coronal Sulcus" = "Foreskin/Coronal Sulcus",
        "Foreskin + Coronal Sulcus" = "Foreskin/Coronal Sulcus",
        "Glans" = "Glans alone",
        "Glans + Coronal Sulcus" = "Glans + Adjacent",
        "Glans + Foreskin" = "Glans + Adjacent",
        "Glans + Shaft" = "Glans + Adjacent"
        ))
Data$pi <- revalue(as.factor(Data$pi), c(
        "2" = "Score 2",
        "3" = "Score 3",
        "4" = "Score 4",
        "5" = "Score 5",
        "6" = "Score 5",
        "7" = "Score 7"
        ))
# Defining colors for plots
nice.colors <- c("turquoise", "coral", "cornflowerblue")
# Defining "density" as 100*met_dens
Data$density <- 100*Data$met_dens
```

***

### Abstract

__Background:__ Proper clinical management of patients with penile squamous cell carcinomas (SCC) is hampered by its rarity in developed nations and by the lack of sufficiently large series with adequate follow-up. Aimed to solve this issue several pathologic features have been proposed as prognostic factors. Recently, nodal density (defined as the ratio of positive inguinal lymph nodes over the total number of inguinal lymph nodes) has been identified as a promising tool to define the outcome of patients with penile SCC. In this study, we evaluate the usefulness of nodal density as a predictor of aggressive pathologic features and outcome in a large series of patients with penile SCC.

__Design:__ This study includes 154 patients with invasive penile SCC treated by penectomy, of whom 133 received a bilateral inguinal lymphadenectomy. The number of lymph nodes per patient that were pathologically evaluated ranged from 1 to 27 for each side, with a median of 8 lymph nodes per side per patient. Clinical endpoints included cancer-free survival and cancer-specific death. Patients were followed  from 11 to 82 months (mean of 54.2 ± 16.5 months). Associations were evaluated using the Mann-Whitney U/Kruskal-Wallis test. A 2-sided P < 0.05 was required for statistical signficance.

__Results:__ High nodal density was significantly associated with the following aggressive pathologic features: invasion of multiple anatomical sites (P=1.36e-4), tumor size > 4 cm (P=1.02e-11), high histologic grade (P=8.36e-13), tumor thickness > 13 mm (P=2.75e-20), presence of lymphovascular invasion (P=1.68e17),  perineural invasion (P=2.48e-9), invasion of deep erectile tissues (P=2.29e-15), prognostic index > 4 (P=2.55e-14), > pT1b (P=2.43e-6), > pN1 (P=1.07e-27), pM (P=4.14e-19), > stage II (P=2.14e-20), and invasion of distal urethra (P=2.35e-6). Regarding outcome, high nodal density was associated with cancer-free survival (P=1.38e-18) and cancer-specific death (P=2.32e-12).

__Conclusions:__ High nodal density is associated with aggressive pathologic features in the primary tumor of patients with penile SCC treated by penectomy and bilateral inguinal lymphadenectomy. High nodal density is significantly associated with a lower proportion of cancer-free survival and a higher proportion of cancer-specific death.

### Results

#### Clinicopathologic Features
* __Patients' age__
```{r DescriptiveAge, fig.width = 9}
descriptive(Data$age)
par(mfrow = c(1,2), cex.lab = 1.25)
# Histogram
hist(Data$age, freq = FALSE, col = "coral", main = "",
     xlab = "Patients' Age (years)", ylab = "Probability Density")
lines(density(Data$age), lwd = 1.5)
# Boxplot
boxplot(Data$age, col = "turquoise", ylab = "Patients' Age (years)")
```

* __Surgical procedure__
```{r DescriptiveSurg}
table.prop(Data$penectomy)
# Barplot
DescriptiveSurg.plot <- barplot(table(Data$penectomy), col = nice.colors,
                                main = "Penectomy", cex.main = 1.5,
                                ylim = c(0, 110), ylab = "No. Patients", cex.lab = 1.25)
text(DescriptiveSurg.plot, paste("N =", table(Data$penectomy)), y = table(Data$penectomy) + 4)
```

* __Anatomical site__
```{r DescriptiveSite}
table.prop(Data$site)
table.prop(Data$site2)
# Barplot
DescriptiveSite.plot <- barplot(table(Data$site2), col = nice.colors,
                                main = "Anatomical Site", cex.main = 1.5,
                                ylim = c(0, 110), ylab = "No. Patients", cex.lab = 1.25)
text(DescriptiveSite.plot, paste("N =", table(Data$site2)), y = table(Data$site2) + 4)
```

* __Tumor size (cm)__
```{r DescriptiveSize, fig.width = 9}
descriptive(Data$size)
par(mfrow = c(1, 2), cex.lab = 1.25)
# Histogram
hist(Data$size, main = "", col = "coral", freq = FALSE,
     xlab = "Tumor size (cm)", ylab = "Probability Density")
lines(density(Data$size), lwd = 1.5)
# Boxplot
boxplot(Data$size, col = "turquoise", ylab = "Tumor size (cm)")
```

* __Tumor subtype__
```{r DescriptiveSubtype, fig.width = 7, fig.height = 7}
table.prop(Data$subtype)
table.prop(Data$subtype2)
SubtypeOrdered <- sort(table(Data$subtype2), decreasing = FALSE)
# Barplot
par(mar = c(5, 9, 0, 2) + 0.1, las = 2, cex.lab = 1.25)
DescriptiveSubtype.plot <- barplot(SubtypeOrdered, horiz = TRUE,
        xlab = "No. Patients",
        col = rainbow(13), xlim = c(0, 80))
text(DescriptiveSubtype.plot, paste("N =", SubtypeOrdered), x = SubtypeOrdered + 5)
```

* __Anatomical level__
```{r DescriptiveLevel, fig.width = 7}
table.prop(Data$level)
# Barplot
DescriptiveLevel.plot <- barplot(table(Data$level), col = nice.colors,
                                 main = "Anatomical Level", cex.main = 1.5,
                                 ylab = "No. Patients", ylim = c(0, 80), cex.lab = 1.25)
text(DescriptiveLevel.plot, paste("N =", table(Data$level)), y = table(Data$level) + 4)
```

* __Histologic grade__
```{r DescriptiveGrade}
table.prop(Data$grade)
# Barplot
DescriptiveGrade.plot <- barplot(table(Data$grade), col = nice.colors,
                                 main = "Histologic Grade", cex.main = 1.5,
                                 ylab = "No. Patients", ylim = c(0, 95), cex.lab = 1.25)
text(DescriptiveGrade.plot, paste("N =", table(Data$grade)), y = table(Data$grade) + 4)
```

* __Tumor thickness__
```{r DescriptiveThickness, fig.width = 9}
descriptive(Data$thick)
par(mfrow = c(1, 2), cex.lab = 1.25)
hist(Data$thick, , freq = FALSE, col = "coral", main = "",
     xlab = "Tumor Thickness (mm)", ylab = "Probability Density", ylim = c(0, 0.06))
lines(density(Data$thick), lwd = 1.5)
boxplot(Data$thick, col = "turquoise", ylab = "Tumor Thickness (mm)")
```

* __Lymphpvascular invasion__
```{r DescriptiveLVI}
table.prop(Data$lvi)
# Barplot
DescriptiveLVI.plot <- barplot(table(Data$lvi), col = nice.colors,
                               main = "Lymphovascular Invasion", cex.main = 1.5,
                               ylab = "No. Patients", ylim = c(0, 100), cex.lab = 1.25)
text(DescriptiveLVI.plot, paste("N =", table(Data$lvi)), y = table(Data$lvi) + 4)
```

* __Perineural invasion__
```{r DescriptivePNI}
table.prop(Data$pni)
# Barplot
DescriptivePNI.plot <- barplot(table(Data$pni), col = nice.colors,
                               main = "Perineural Invasion", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 110), cex.lab = 1.25)
text(DescriptivePNI.plot, paste("N =", table(Data$pni)), y = table(Data$pni) + 4)
```

* __HPV infection__
```{r DescriptiveHPV}
table.prop(Data$hpvrelated)
# Barplot
DescriptiveHPV.plot <- barplot(table(Data$hpvrelated), col = nice.colors,
                               main = "HPV Infection", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 110), cex.lab = 1.25)
text(DescriptiveHPV.plot, paste("N =", table(Data$hpvrelated)), y = table(Data$hpvrelated) + 4)
```

* __Prognostic index__
```{r DescriptivePI}
table.prop(Data$pi)
# Barplot
DescriptivePI.plot <- barplot(table(Data$pi), col = rainbow(6),
                               main = "Prognostic Index", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 55), cex.lab = 1.25)
text(DescriptivePI.plot, paste("N =", table(Data$pi)), y = table(Data$pi) + 4)
```

* __pT stage__
```{r DescriptiveT}
table.prop(Data$pt_stage)
# Barplot
DescriptiveT.plot <- barplot(table(Data$pt_stage), col = rainbow(5),
                               main = "pT Stage", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 100), cex.lab = 1.25)
text(DescriptiveT.plot, paste("N =", table(Data$pt_stage)), y = table(Data$pt_stage) + 4)
```

* __pN stage__
```{r DescriptiveN}
table.prop(Data$pn_stage)
# Barplot
DescriptiveN.plot <- barplot(table(Data$pn_stage), col = nice.colors,
                               main = "pN Stage", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 100), cex.lab = 1.25)
text(DescriptiveN.plot, paste("N =", table(Data$pn_stage)), y = table(Data$pn_stage) + 4)
```

* __pM stage__
```{r DescriptiveM}
table.prop(Data$m_stage)
# Barplot
DescriptiveM.plot <- barplot(table(Data$m_stage), col = nice.colors,
                               main = "M Stage", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 100), cex.lab = 1.25)
text(DescriptiveM.plot, paste("N =", table(Data$m_stage)), y = table(Data$m_stage) + 4)
```

* __TNM stage__
```{r DescriptiveTNM}
table.prop(Data$tnm_stage)
# Barplot
DescriptiveTNM.plot <- barplot(table(Data$tnm_stage), col = rainbow(5),
                               main = "TNM Stage", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 100), cex.lab = 1.25)
text(DescriptiveTNM.plot, paste("N =", table(Data$tnm_stage)), y = table(Data$tnm_stage) + 4)
```

* __Urethral invasion__
```{r DescriptiveU}
table.prop(Data$urethra)
# Barplot
DescriptiveU.plot <- barplot(table(Data$urethra), col = nice.colors,
                               main = "Urethral Invasion", cex.main = 1.5, 
                               ylab = "No. Patients", ylim = c(0, 120), cex.lab = 1.25)
text(DescriptiveU.plot, paste("N =", table(Data$urethra)), y = table(Data$urethra) + 4)
```

#### Nodal Density
* Diagnosis of lymphadenopathy
```{r NodesDx}
table.prop(Data$ln_dx)
DataNodes <- subset(Data, ln_dx == "Lymphadenectomy")
```

* Lymph nodes metastasis
```{r NodesMetastasis}
table.prop(DataNodes$ln_met)
table.prop(DataNodes$met_bi)
```

* Right lymph nodes
```{r NodesRight}
descriptive(DataNodes$ln_r)
descriptive(DataNodes$r_met)
```

* Left lymph nodes
```{r NodesLeft}
descriptive(DataNodes$ln_l)
descriptive(DataNodes$l_met)
```

#### Association of Nodal Density with Clinicopathologic Features
* __Nodal Density and Anatomical Site of Primary Tumor__
```{r NodalSite, fig.width = 9}
compare.kw(DataNodes$density, DataNodes$site2)
# Significance test
NodalSite.P <- kruskal.test(density ~ site2, data = DataNodes)
# Boxplot
boxplot(density ~ site2, data = DataNodes, cex.lab = 1.25,
        main = "Nodal Density and Anatomical Site of Primary Tumor", cex.main = 1.5,
        names = c("Foreskin + Coronal Sulcus", "Glans Alone", "Glans + Adjacent"),
        col = nice.colors,
        ylab = "Nodal Density")
text(paste("P value =", format(NodalSite.P$p.value, digits = 3, width = 7, scientific = TRUE)), x = 0.85, y = 85)
```

* __Nodal density and tumor size__
```{r NodalSize, fig.width = 9, fig.height = 7}
NodalSize.Cor <- cor.test(DataNodes$density, DataNodes$size, method = c("spearman"))
NodalSize.Cor
# Scatterplot
scatterplot(density ~ size, data = DataNodes, boxplots = FALSE, grid = FALSE,
            main = "Nodal Density and Tumor Size", cex.main = 1.5,
            ylab = "Nodal Density", xlab = "Tumor Size (cm)", cex.lab = 1.25)
text(paste("P value =", format(NodalSize.Cor$p.value, digits = 3, width = 7)), x = 2, y = 85)
text(paste("Spearman's rho =", format(NodalSize.Cor$estimate, digits = 3, width = 7)), x = 2.3, y = 77.5)
```

* __Nodal density and histologic subtype__
```{r NodalSubtype, fig.width = 9, fig.height = 7}
compare.kw(DataNodes$density, DataNodes$subtype2)
# Significance test
NodalSubtype.P <- kruskal.test(density ~ subtype2, data = DataNodes)
# Boxplot
par(mar = c(5, 9, 4, 2) + 0.1, las = 2, cex.lab = 1.25)
boxplot(density ~ subtype2, data = DataNodes, horizontal = TRUE,
        main = "Nodal Density and Histologic Subtype", cex.main = 1.5,
        col = rainbow(13), xlab = "Nodal Density")
text(paste("P value =", format(NodalSubtype.P$p.value, digits = 3, width = 7, scientific = TRUE)), x = 80, y = 13)
```

* __Nodal density and histologic grade__
```{r NodalGrade, fig.width = 9}
compare.kw(DataNodes$density, DataNodes$grade)
# Significance test
NodalGrade.P <- kruskal.test(density ~ grade, data = DataNodes)
# Boxplot
boxplot(density ~ grade, data = DataNodes, cex.lab = 1.25,
        main = "Nodal Density and Histologic Grade", cex.main = 1.5,
        col = nice.colors, ylab = "Nodal Density")
text(paste("P value =", format(NodalGrade.P$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
```

* __Nodal density and tumor thickness__
```{r NodalThickness, fig.width = 9, fig.height = 7}
# Significance test
NodalThickness.Cor <- cor.test(DataNodes$density, DataNodes$thick, method = c("spearman"))
NodalThickness.Cor
# Scatterplot
scatterplot(density ~ thick, data = DataNodes, boxplots = FALSE, grid = FALSE,
            main = "Nodal Density and Tumor Thickness", cex.main = 1.5,
            xlab = "Tumor Thickness (mm)", ylab = "Nodal Density", cex.lab = 1.25)
text(paste("P value =", format(NodalThickness.Cor$p.value, digits = 3, width = 7)), x = 4.5, y = 85)
text(paste("Spearman's rho =", format(NodalThickness.Cor$estimate, digits = 3, width = 7)), x =5.25, y = 77.5)

```

* __Nodal density and lymphovascular invasion__
```{r NodalLVI, fig.width = 7}
compare.mw(DataNodes$density, DataNodes$lvi)
# Significance test
NodalLVI.P <- wilcox.test(density ~ lvi, data = DataNodes)
# Boxplot
boxplot(density ~ lvi, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and Lymphovascular Invasion", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Lymphovascular Invasion")
text(paste("P value =", format(NodalLVI.P$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
```

* __Nodal density and perineural invasion__
```{r NodalPNI, fig.width = 7}
compare.mw(DataNodes$density, DataNodes$pni)
# Significance test
NodalPNI.P <- wilcox.test(density ~ pni, data = DataNodes)
boxplot(density ~ pni, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and Perineural Invasion", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Perineural Invasion")
text(paste("P value =", format(NodalPNI.P$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
```

* __Nodal density and anatomical level__
```{r NodalLevel, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$level)
# Significance test
NodalLevel.P <- kruskal.test(density ~ level, data = DataNodes)
# Boxplot
boxplot(density ~ level, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and Anatomical Level", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Anatomical Level")
text(paste("P value =", format(NodalLevel.P$p.value, digits = 3, width = 7)), x = 1, y = 85)
```

* __Nodal density and HPV infection__
```{r NodalHPV, fig.width = 7}
compare.mw(DataNodes$density, DataNodes$hpvrelated)
# Significance test
NodalHPV.P <- kruskal.test(density ~ hpvrelated, data = DataNodes)
# Boxplot
boxplot(density ~ hpvrelated, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and HPV Infection", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "HPV Infection")
text(paste("P value =", format(NodalHPV.P$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
```

* __Nodal density and prognostic index__
```{r NodalPI, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$pi)
# Significance test
NodalPI.P <- kruskal.test(density ~ pi, data = DataNodes)
# Boxplot
boxplot(density ~ pi, data = DataNodes, cex.lab = 1.25, col = rainbow(6),
        main = "Nodal Density and Prognostic Index", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Prognostic Index")
text(paste("P value =", format(NodalPI.P$p.value, digits = 3, width = 7)), x = 1.25, y = 85)
```

* __Nodal density and pT stage__
```{r NodalT, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$pt_stage)
# Significance test
NodalT.P <- kruskal.test(density ~ pt_stage, data = DataNodes)
# Boxplot
boxplot(density ~ pt_stage, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and pT Stage", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "pT Stage")
text(paste("P value =", format(NodalT.P$p.value, digits = 3, width = 7)), x = 1.25, y = 85)
```

* __Nodal density and pN stage__
```{r NodalN, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$pn_stage)
# Significance test
NodalN.P <- kruskal.test(density ~ pn_stage, data = DataNodes)
# Boxplot
boxplot(density ~ pn_stage, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and pN Stage", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "pN Stage")
text(paste("P value =", format(NodalN.P$p.value, digits = 3, width = 7)), x = 1, y = 85)
```

* __Nodal density and pM stage__
```{r NodalM, fig.width = 7}
compare.mw(DataNodes$density, DataNodes$m_stage)
# Significance test
NodalM.P <- wilcox.test(density ~ m_stage, data = DataNodes)
# Boxplot
boxplot(density ~ m_stage, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and pM Stage", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "pM Stage")
text(paste("P value =", format(NodalM.P$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
```

* __Nodal density and TNM stage__
```{r NodalTNM, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$tnm_stage)
# Significance test
NodalTNM.P <- kruskal.test(density ~ tnm_stage, data = DataNodes)
# Boxplot
boxplot(density ~ tnm_stage, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and TNM Stage", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "TNM Stage")
text(paste("P value =", format(NodalTNM.P$p.value, digits = 3, width = 7)), x = 1.5, y = 85)
```

* __Nodal density and urethral invasion__
```{r NodalUrethra, fig.width = 7}
compare.mw(DataNodes$density, DataNodes$urethra)
# Significance test
NodalUrethra.P <- wilcox.test(density ~ urethra, data = DataNodes)
# Boxplot
boxplot(density ~ urethra, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and Urethral Invasion", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Urethral Invasion")
text(paste("P value =", format(NodalUrethra.P$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
```

#### Association of Nodal Density with Outcome
* __Distribution of outcomes__
```{r NodalOutcome}
table.prop(DataNodes$outcome)
```
* __Follow-up lenght__
```{r NodalFU}
descriptive(DataNodes$fu_month)
```

* __Nodal density and cancer-disease (AWD & DOD)__
```{r NodalDisease, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$cancer_dis)
# Kruskal-Wallis test
NodalDiseaseKW <- kruskal.test(density ~ cancer_dis, data = DataNodes)
# Cohen's d test
NodalDiseaseCohen <- cohen.d(density ~ cancer_dis, data = DataNodes)
# Boxplot
boxplot(density ~ cancer_dis, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and Cancer-Disease", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Cancer-Disease")
text(paste("P value =", format(NodalDiseaseKW$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
text(paste("Cohen's d =", format(NodalDiseaseCohen$estimate, digits = 3, width = 7)), x = 0.75, y = 77.5)
```

* __Nodal density and cancer-death__
```{r NodalDOD, fig.width = 7}
compare.kw(DataNodes$density, DataNodes$cancer_death)
# Kruskal-Wallis test
NodalDODKW <- kruskal.test(density ~ cancer_death, data = DataNodes)
# Cohen's d test
NodalDODCohen <- cohen.d(density ~ cancer_death, data = DataNodes)
# Boxplot
boxplot(density ~ cancer_death, data = DataNodes, cex.lab = 1.25, col = nice.colors,
        main = "Nodal Density and Cancer-Death", cex.main = 1.5,
        ylab = "Nodal Density", xlab = "Cancer-Death")
text(paste("P value =", format(NodalDODKW$p.value, digits = 3, width = 7)), x = 0.75, y = 85)
text(paste("Cohen's d =", format(NodalDODCohen$estimate, digits = 3, width = 7)), x = 0.75, y = 77.5)
```
